---
title: "Week 10 Final Project Work"
author: "Wong Zi Xin"
date: "`r Sys.Date()`"
output: html_document
---

library(shiny)
library(ggplot2)
library(plotly)
library(dplyr)
library(shinydashboard)
library(reshape2)
library(scales)

# Define UI
ui <- fluidPage(
  
  # App title
  titlePanel("Data Story: The Big Picture"),
  
  # Create tabs for different graphs
  tabsetPanel(
    tabPanel("Number of Passengers by Country",
             # Output: Plotly Barplot
             plotlyOutput(outputId = "barPlot1")
    ),
    tabPanel("Number of Passengers by Year",
             # Output: Plotly Barplot
             plotlyOutput(outputId = "barPlot2")
    ),
    tabPanel("Heatmap of Passengers",
             # Output: Heatmap
             plotOutput(outputId = "heatmap")
    )
  )
)

# Define server logic
server <- function(input, output) { 
  
  # Read the CSV data into data frames
  data_country <- read.csv("country_passengers.csv")
  data_year <- read.csv("year_passengers.csv")
  data_heatmap <- read.csv("heatmap.csv", skip = 1, header = TRUE)
  
  # Melt the heatmap data, reorder it in alphabetical order from top to bottom
  data_melted <- melt(data_heatmap, id.vars = "Country", variable.name = "Year", value.name = "Value")
  data_melted$Year <- gsub("^X", "", data_melted$Year)
  data_melted$Country <- factor(data_melted$Country, levels = rev(sort(unique(data_melted$Country))))
  
  # Render the first graph
  output$barPlot1 <- renderPlotly({
    
    # Create a ggplot object
    p <- ggplot(data_country, aes(x = Country, y = Number.of.Passengers, text = paste("Country: ", Country, "<br>Passengers: ", scales::comma(Number.of.Passengers)))) +
      geom_bar(stat = "identity", fill = "grey", color = "white") +
      labs(x = "Country", y = "Number of Passengers", title = "Number of Passengers by Country") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_y_continuous(labels = scales::comma, breaks = seq(0, max(data_country$Number.of.Passengers), by = 200000000))  # Customize y-axis breaks
    
    # Convert ggplot object to a plotly object
    ggplotly(p, tooltip = "text")
    
  })
  
  # Render the second graph
  output$barPlot2 <- renderPlotly({
    
    # Create a ggplot object
    p <- ggplot(data_year, aes(x = Year, y = Number.of.Passengers, text = paste("Year: ", Year, "<br>Passengers: ", scales::comma(Number.of.Passengers)))) +
      geom_bar(stat = "identity", fill = "grey", color = "white") +
      labs(x = "Year", y = "Number of Passengers", title = "Number of Passengers by Year") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_y_continuous(labels = scales::comma, breaks = seq(0, max(data_year$Number.of.Passengers), by = 100000000))  # Customize y-axis breaks
    
    # Convert ggplot object to a plotly object
    ggplotly(p, tooltip = "text")
    
  })
  
  # Render the heatmap
  output$heatmap <- renderPlot({
    ggplot(data_melted, aes(x = Year, y = Country, fill = Value)) + 
      geom_tile() + 
      scale_fill_gradient(low = "white", high = "red", labels = scales::comma) +
      theme_minimal() +
      labs(title = "Heatmap", x = "Year", y = "Country")
    
  })
  
}

# Create Shiny app
shinyApp(ui = ui, server = server)




library(shiny)
library(plotly)
library(shinydashboard)
library(scales)
library(tidyverse)

#Creating a new column for replacing 'geo' values 
country_mapping <- c(
  AT = "Austria", BE = "Belgium", BG = "Bulgaria", CY = "Cyprus",
  CZ = "Czechia", DE = "Germany", DK = "Denmark", EE = "Estonia",
  EL = "Greece", ES = "Spain", FI = "Finland", FR = "France",
  HR = "Croatia", HU = "Hungary", IE = "Ireland", IT = "Italy",
  LT = "Lithuania", LU = "Luxembourg", LV = "Latvia", MT = "Malta",
  NL = "Netherlands", PL = "Poland", PT = "Portugal", RO = "Romania",
  SE = "Sweden", SI = "Slovenia", SK = "Slovakia"
)

# Define UI
ui <- fluidPage(
  titlePanel("Data Story: The Big Picture"),
  tabsetPanel(
    tabPanel("Number of Passengers by Country", plotlyOutput(outputId = "barPlot1")),
    tabPanel("Number of Passengers by Year", plotlyOutput(outputId = "barPlot2")),
    tabPanel("Heatmap of Passengers", plotOutput(outputId = "heatmap"))
  )
)

# Define server logic
server <- function(input, output) { 
  # Using read.csv to read the data
  data <- read.csv("EU27_passenger_data.csv")
  
  # Clean the data
  cleaned_data <- data %>%
    select(geo, TIME_PERIOD, OBS_VALUE) %>%          # Select required columns
    pivot_wider(names_from = TIME_PERIOD,            # Spread the data
                values_from = OBS_VALUE) %>%
    #Replace country codes with country names in new column
    mutate(CountryName = ifelse(geo %in% names(country_mapping), country_mapping[geo], geo))
  
  # List of countries to exclude
  countries_to_exclude <- c("BA", "CH", "EU27_2020", "IS", "ME", "MK", "NO", "RS", "UK")
  
  # Exclude the countries from the dataset
  filtered_data <- cleaned_data %>%
    filter(!(geo %in% countries_to_exclude))
  
  # Melt the data for the heatmap and rename the 'geo' column to 'Country'
  data_melted <- filtered_data %>%
    pivot_longer(cols = -c(geo, CountryName), names_to = "Year", values_to = "Value") %>%
    rename(Country = CountryName) %>% 
    select(-geo) #Drop the 'geo' column
  
  # Reorder countries in alphabetical order from top to bottom
  data_melted$Country <- factor(data_melted$Country, levels = rev(sort(unique(data_melted$Country))))
  
  # Prepare datasets for plots
  data_country <- filtered_data %>% select(CountryName, `2020`) %>% rename(Country = CountryName, Number.of.Passengers = `2020`)
  data_year <- data_melted %>% group_by(Year) %>% summarise(Number.of.Passengers = sum(Value, na.rm = TRUE))
  
  # Render the first graph
  output$barPlot1 <- renderPlotly({
    p <- ggplot(data_country, aes(x = Country, y = Number.of.Passengers, text = paste("Country: ", Country, "<br>Passengers: ", scales::comma(Number.of.Passengers)))) +
      geom_bar(stat = "identity", fill = "grey", color = "white") +
      labs(x = "Country", y = "Number of Passengers", title = "Number of Passengers by Country from 2011 to 2022") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_y_continuous(labels = scales::comma, breaks = seq(0, max(data_country$Number.of.Passengers), by = 5000000))
    ggplotly(p, tooltip = "text")
  })
  
  # Render the second graph
  output$barPlot2 <- renderPlotly({
    p <- ggplot(data_year, aes(x = Year, y = Number.of.Passengers, text = paste("Year: ", Year, "<br>Passengers: ", scales::comma(Number.of.Passengers)))) +
      geom_bar(stat = "identity", fill = "grey", color = "white") +
      labs(x = "Year", y = "Number of Passengers", title = "Number of Passengers by Year") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_y_continuous(labels = scales::comma, breaks = seq(0, max(data_year$Number.of.Passengers), by = 100000000))
    ggplotly(p, tooltip = "text")
  })
  
  # Render the heatmap
  output$heatmap <- renderPlot({
    ggplot(data_melted, aes(x = Year, y = Country, fill = Value)) + 
      geom_tile() + 
      scale_fill_gradient(low = "white", high = "red", labels = scales::comma) +
      theme_minimal() +
      labs(title = "Heatmap", x = "Year", y = "Country")
  })
}

# Create Shiny app
shinyApp(ui = ui, server = server)

```{r, eval=TRUE,echo=TRUE}
library(ggplot2)
library(tidyverse)
```

```{r, eval=TRUE,echo=TRUE}
data <- read.csv("EU27_passenger_data.csv")
```

```{r, eval=TRUE,echo=TRUE}
# Your previous data cleaning steps
cleaned_data <- data %>%
    select(geo, TIME_PERIOD, OBS_VALUE) %>%
    pivot_wider(names_from = TIME_PERIOD, values_from = OBS_VALUE) %>%
    mutate(CountryName = ifelse(geo %in% names(country_mapping), country_mapping[geo], geo))

# Filter the data for years 2019 and 2020
scatter_data <- cleaned_data %>%
    select(CountryName, `2019`, `2020`)

ggplot(scatter_data, aes(x = `2019`, y = `2020`, label = CountryName)) +
  geom_point(aes(color = CountryName), size = 3) +
  geom_text(aes(label = CountryName), vjust = 2, hjust = 2, size = 3) +
  labs(
    x = "Number of Passengers (2019)",
    y = "Number of Passengers (2020)",
    title = "Comparison of Air Passenger Numbers: 2019 vs 2020",
    subtitle = "Each point represents a country",
    color = "Country"
  ) +
  theme_minimal()
```

library(shiny)
library(plotly)
library(shinydashboard)
library(scales)
library(tidyverse)

# Creating a new column for replacing 'geo' values
country_mapping <- c(
  AT = "Austria", BE = "Belgium", BG = "Bulgaria", CY = "Cyprus",
  CZ = "Czechia", DE = "Germany", DK = "Denmark", EE = "Estonia",
  EL = "Greece", ES = "Spain", FI = "Finland", FR = "France",
  HR = "Croatia", HU = "Hungary", IE = "Ireland", IT = "Italy",
  LT = "Lithuania", LU = "Luxembourg", LV = "Latvia", MT = "Malta",
  NL = "Netherlands", PL = "Poland", PT = "Portugal", RO = "Romania",
  SE = "Sweden", SI = "Slovenia", SK = "Slovakia"
)

# Define UI
ui <- fluidPage(
  titlePanel("Data Story: The Big Picture"),
  tabsetPanel(
    tabPanel("Number of Passengers by Country", plotlyOutput(outputId = "barPlot1")),
    tabPanel("Number of Passengers by Year", plotlyOutput(outputId = "barPlot2")),
    tabPanel("Heatmap of Passengers", plotOutput(outputId = "heatmap")),
    tabPanel("Passengers Over Years by Country", 
             uiOutput("selectCountry"), # This will get the updated choices
             plotlyOutput(outputId = "barPlot3"))
    tabPanel("Air Passenger Numbers: 2019 vs 2020", plotlyOutput("scatterPlot"))
  )
)

# Define server logic
server <- function(input, output, session) { 
  # Using read.csv to read the data
  data <- read.csv("EU27_passenger_data.csv")
  
  # Clean the data
  cleaned_data <- data %>%
    select(geo, TIME_PERIOD, OBS_VALUE) %>%          # Select required columns
    pivot_wider(names_from = TIME_PERIOD,            # Spread the data
                values_from = OBS_VALUE) %>%
    mutate(CountryName = ifelse(geo %in% names(country_mapping), country_mapping[geo], geo)) # Replace country codes
  
  # List of countries to exclude
  countries_to_exclude <- c("BA", "CH", "EU27_2020", "IS", "ME", "MK", "NO", "RS", "UK")
  
  # Exclude the countries from the dataset
  filtered_data <- cleaned_data %>%
    filter(!(geo %in% countries_to_exclude))
  
  # Melt the data for the heatmap and rename the 'geo' column to 'Country'
  data_melted <- filtered_data %>%
    pivot_longer(cols = -c(geo, CountryName), names_to = "Year", values_to = "Value") %>%
    rename(Country = CountryName) %>% 
    select(-geo) #Drop the 'geo' column
  
  # Reorder countries in alphabetical order from top to bottom
  data_melted$Country <- factor(data_melted$Country, levels = rev(sort(unique(data_melted$Country))))
  
  # Prepare datasets for plots
  data_country <- filtered_data %>% select(CountryName, `2020`) %>% rename(Country = CountryName, Number.of.Passengers = `2020`)
  data_year <- data_melted %>% group_by(Year) %>% summarise(Number.of.Passengers = sum(Value, na.rm = TRUE))
  
  # Extracting 2019 and 2020 data for scatter plot
  scatter_data <- cleaned_data %>%
    select(Country, `2019`, `2020`) %>%
    gather(key = "Year", value = "Passengers", -Country) %>%
    spread(key = Year, value = Passengers)
  
  # Render the first graph
  output$barPlot1 <- renderPlotly({
    p <- ggplot(data_country, aes(x = Country, y = Number.of.Passengers, text = paste("Country: ", Country, "<br>Passengers: ", scales::comma(Number.of.Passengers)))) +
      geom_bar(stat = "identity", fill = "grey", color = "white") +
      labs(x = "Country", y = "Number of Passengers", title = "Number of Passengers by Country") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_y_continuous(labels = scales::comma, breaks = seq(0, max(data_country$Number.of.Passengers), by = 5000000))
    ggplotly(p, tooltip = "text")
  })
  
  # Render the second graph
  output$barPlot2 <- renderPlotly({
    p <- ggplot(data_year, aes(x = Year, y = Number.of.Passengers, text = paste("Year: ", Year, "<br>Passengers: ", scales::comma(Number.of.Passengers)))) +
      geom_bar(stat = "identity", fill = "grey", color = "white") +
      labs(x = "Year", y = "Number of Passengers", title = "Number of Passengers by Year") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_y_continuous(labels = scales::comma, breaks = seq(0, max(data_year$Number.of.Passengers), by = 100000000))
    ggplotly(p, tooltip = "text")
  })
  
  # Render the heatmap
  output$heatmap <- renderPlot({
    ggplot(data_melted, aes(x = Year, y = Country, fill = Value)) + 
      geom_tile() + 
      scale_fill_gradient(low = "white", high = "red", labels = scales::comma) +
      theme_minimal() +
      labs(title = "Heatmap", x = "Year", y = "Country")
  })
  
  # Ensure the selectInput is updated after the data is prepared
  output$selectCountry <- renderUI({
    selectInput("selectedCountry", "Select a Country:", choices = unique(data_melted$Country))
  })
  
  # Render the third graph
  output$barPlot3 <- renderPlotly({
    req(input$selectedCountry)  # Ensure that input$selectedCountry is available
    country_data <- data_melted %>%
      filter(Country == input$selectedCountry) %>%
      mutate(Year = as.numeric(Year), Value = as.numeric(Value)) %>%
      arrange(Year)  # Sort the data by Year
    
    p <- ggplot(country_data, aes(x = Year, y = Value, group = 1, text = paste("Year: ", Year, "<br>Passengers: ", Value))) +
      geom_line() +
      geom_point() +
      labs(x = "Year", y = "Number of Passengers", title = paste("Number of Passengers over the Years -", input$selectedCountry)) +
      scale_x_continuous(breaks = 2011:2022) +
      theme_minimal()
    
    ggplotly(p, tooltip = "text")
  })
  
  # Render the scatter plot
  output$scatterPlot <- renderPlotly({
    p <- ggplot(scatter_data, aes(x = `2019`, y = `2020`, text = Country)) +
      geom_point(size = 3) +
      geom_smooth(method = "lm", se = FALSE, color = "grey") +
      labs(title = "Air Passenger Numbers: 2019 vs 2020",
           x = "2019",
           y = "2020") +
      theme_minimal()
    
    ggplotly(p, tooltip = "text")
  })
  
}

# Create Shiny app
shinyApp(ui = ui, server = server)